---
title: "Lab 1"
output:
  html_document: default
  pdf_document: default
---

### Enter the names of the group members below: (2 pts)

**This assignment is due by the end of the day. Only one student in the group submits a pdf file on Gradescope. When submitting the lab, remember to ADD GROUP MEMBERS to the submission.**

*For all questions, include the R commands/functions that you used to find your answer (show R chunk). Answers without supporting code will not receive credit. Write full sentences to describe your findings.*

In this lab, you will explore the dataset `faithful`. It contains information about eruptions of the Old Faithful geyser in Yellowstone National Park. The first few observations are listed below.

```{r, warning=FALSE,message=FALSE}
## Load the R packages that you will need for this Lab

library(tidyverse)
library(datasets)
faithful |> 
    head()
```

### Question 1: (2 pts)

The first goal is to determine how many rows and columns are in the dataset. Without looking at the data, have each group member guess how how many rows there are.

**Your answer goes here. Write sentences in bold.**

Try using the `glimpse()` function from the `tidyverse` on the `faithful` dataset. How many rows and columns does the `glimpse()` function report for the `faithful` dataset? Which lab member came the closest to what `glimpse()` reports?

```{r}
# your code goes below (make sure to edit comment)

```

**Your answer goes here. Write sentences in bold.**

In which units are the variables reported? You will need more information about the dataset to answer that last question: run `?faithful` **in the console**. 

*Note: DO NOT use `?` in this document as that will prevent you from rendering it or will open the documentation in a new tab.*

**Your answer goes here. Write sentences in bold.**

------------------------------------------------------------------------

### Question 2: (2 pts)

Using the function `summary()` for each variable, find the minimum, maximum, mean, and median values of each variable. Write a sentence to interpret the mean eruption duration and the mean waiting time.

```{r}
# your code goes below (make sure to edit comment)

```


**Your answer goes here. Write sentences in bold.**

Two scientists studying the waiting times between eruptions at Old Faithful have different theories of the mean waiting time between eruptions. Scientist A hypothesizes that the true mean waiting time is 50 minutes while Scientist B hypothesizes that the true mean waiting time is 75 minutes. 

When considering the output of the `summary()` function, which hypothesis seems more likely according to the data (A or B)?

**Your answer goes here. Write sentences in bold.**

------------------------------------------------------------------------

### Question 3: (2 pts)

Create a new variable with the function `cut_number()` along with `mutate()`. (You can find the help page for `cut_number()` by calling `?cut_number` at the console.) 


The code below divides the `eruptions` variable into 5 categories and stores the categories in a new variable called `eruptions_cat`. Modify the code below to create a tabulation of the `eruptions` variable using 10 categories.

```{r}
# your code goes below (make sure to edit comment)
faithful |> 
    mutate(eruptions_cat = cut_number(eruptions, 5)) 
```


Extend the code above to divide the `waiting` variable into 8 categories and create a new variable called `waiting_cat`.

```{r}
# your code goes below (make sure to edit comment)

```



------------------------------------------------------------------------

### Question 4: (2 pts)

Let's do some filtering and logical indexing. What does the code below do?

```{r, eval=FALSE}
faithful |> 
        filter(waiting > 60) 
```

**Your answer goes here. Write sentences in bold.**

------------------------------------------------------------------------

### Question 5: (2 pts)

Using filtering and logical indexing and the function `mean()`, find

-   the mean of the variable `eruptions` when `waiting` is **less than or equal to** 1 hour, and

-   the mean of the variable `eruptions` when `waiting` is **more than** 1 hour.

Compare the two means. How do the means change with the waiting time being less than or equal to or greater than 1 hour?

```{r}
# your code goes below (make sure to edit comment)

```

**Your answer goes here. Write sentences in bold.**

------------------------------------------------------------------------

### Question 6: (2 pts)

Use tabulation techniques to explore how the waiting time might be related to the eruption duration eruption time. Divide the `waiting` variable into 10 categories using `cut_number()` and compute the mean of the `eruptions` variable within each category using, `group_by()`, `summarize()` and `mean()`. 

Describe what you see in the tabulation output. Does your interpretation of the tabulation change if you change the number of categories to be more or less than 10?

```{r}
# your code goes below (make sure to edit comment)

```

**Your answer goes here. Write sentences in bold.**



### Question 7: (2 pts)

Consider the following hypotheses about the relationship between waiting time and eruption length:

- A longer waiting time after one eruption is generally followed by a longer eruption compared to an eruption that follows a shorter waiting time;

- The waiting time between eruptions and the eruption length have no relationship at all. 

Given the data shown in the tabulation from Question 6, which of the two hypotheses is more likely?


**Your answer goes here. Write sentences in bold.**

------------------------------------------------------------------------

### Formatting: (4 pts)

Render your file! You can

1. Render your file into HTML by running the Knit command in RStudio (clicking the "Knit" button above the editor)

2. Once it renders in HTML, click on `Open in Browser` at the top left of the window that pops out. (Alternatively, you can open the HTML file in your browser directly.)

3. Print your HTML file into PDF from your browser. 

4. Upload your PDF file to Gradescope.

5. Remember to add your **group members** to the assignment submission.

6. Remember to mark the pages that correspond to each question!
